---
title: "Supplemental Lipid Figures"
author: "Hanna Choi"
date: "11/17/2020"
output:
  html_document:
    code_folding: hide
---
<center>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
library(readr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(DT)
library(ggpubr)
```
# Making usable data for ggplot

```{r data with counts}
POS <- read.csv("Blanchard_Lipid_POS.csv", header=TRUE)
NEG <- read.csv("Blanchard_Lipid_NEG.csv", header=TRUE)
names(POS) <- gsub(x = names(POS), pattern = "_POS", replacement = "")  
names(NEG) <- gsub(x=names(POS), pattern="_NEG", replacement="")
POS$charge <- "POS"
NEG$charge <- "NEG"
data <- rbind(POS, NEG) # lipid data with counts
```
```{r metadata of only significant, message=FALSE}
summary_class <- read_csv("significantlipidswithformula.csv")[-1] # made in comparealigned .Rmd
summary_class <- summary_class %>%
  rename(Main_Class = "Main Class") #getting rid of the space in column
summary_class <- summary_class[order(summary_class$Main_Class, summary_class$Standardized_Difference_Control_minus_Heated),] #ordering the metadata
```
```{r workable dataframe df for ggplot}
df <- merge(summary_class, data, by="row.identity", all=FALSE)
# For the dflong below:
# columns 17-43 is counts 
# Blanch_Nat_Lip is redundant
# separate by treatment Control vs Heated
dflong <- df[-c(11:16)] %>% 
  gather(plot, count,Blanch_Nat_Lip_C_12_AB_M_17:Blanch_Nat_Lip_H_4_AB_M_05) %>% 
  mutate(plot = gsub("Blanch_Nat_Lip_", "", plot)) %>% 
  separate(plot, c("treatment","plot"), "_", extra = "merge") %>%
  unite("row.id", c(row.identity, soil_type), remove=FALSE) %>%
  mutate(row.id = gsub("_Mineral", "", row.id)) %>%
  mutate(row.id = gsub("_Organic", "", row.id)) %>%
  arrange(soil_type)
dflongMineral <- dflong[which(dflong$soil_type == "Mineral"),]
dflongOrganic <- dflong[which(dflong$soil_type == "Organic"),]
```

# Overarching Lipid Figure
```{r slightly messy but all significant lipids ordered by class, fig.cap = "Lipidomic data found 31 lipids that were altered in abundance significantly. 1 of the 31 lipid abundances were found to have increased due to an increase in temperature. From the figure, TG is Triacylglyerol"}
lipidfigureMineral <- ggplot(data=dflongMineral, aes(x=row.id, y=log(count), group_by(Main_Class, treatment), fill = treatment)) + geom_boxplot(show.legend = FALSE) + coord_flip() + scale_fill_manual(values = c( "#082BEA", "#EA0D08")) + scale_color_manual(values = c("#D4E126", "#12EF3A")) + ggtitle("Significant Lipids Heated v. Control Soil Mineral Plots") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
lipidfigureOrganic <- ggplot(data=dflongOrganic, aes(x=row.id, y=log(count), group_by(Main_Class, treatment), fill = treatment)) + geom_boxplot() + coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_fill_manual(values = c( "#082BEA", "#EA0D08")) + scale_color_manual(values = c("#D4E126", "#12EF3A")) + ggtitle("Significant Lipids Heated v. Control Soil Organic Plots")
lipidfigure <- ggarrange(lipidfigureMineral, lipidfigureOrganic, ncol=1)
lipidfigure
```
# Making plots by Class

```{r ggplots}
list.ggplotsbygroup <- dflong %>% group_by(Main_Class) %>% do(plots=ggplot(data=.) + aes(x=row.id, y=log(count)) + geom_boxplot(aes(fill=treatment)) + xlab("compound") + ylab("abundance (logged)") + ggtitle(unique(.$Main_Class)))
```
# Analysis by Group
```{r, fig.cap="DG; inconclusive due to number of compounds being low and no significant differences between chain lengths or degrees of saturation"}
DG1 <- list.ggplotsbygroup$plots[[1]] + coord_flip()
DG1
```

---

```{r, fig.cap = "PC; there's an interesting difference between PC_A and PC_B which only differ in stereochemistry"}
PC2 <- list.ggplotsbygroup$plots[[2]] + coord_flip()
PC2
```

---

```{r, fig.cap = "PE is found to be a plant growth regulator"}
PE3 <- list.ggplotsbygroup$plots[[3]] + coord_flip()
PE3
```

---

```{r, fig.cap = "DGDG; decreased abundance for all compunds"}
DGDG4 <- list.ggplotsbygroup$plots[[4]] + coord_flip()
DGDG4
```

---

```{r, fig.cap = "Only 1 hexosylceramides with two different stereochemical forms found to be significantly altered due to Temeprature. The abundance signficantly decreased in the Heated plots"}
HexCer5 <- list.ggplotsbygroup$plots[[5]] + coord_flip()
HexCer5
```

---

```{r, fig.cap = "Significant DGTSA's decreased in abundance as well"}
DGTSA6 <- list.ggplotsbygroup$plots[[6]] + coord_flip()
DGTSA6
```

---

```{r, fig.cap = "TG's are most common lipids in soil and TG(58:9) was found to be the only lipid to increase in abundance in the heated plots over the control plots"}
TG7 <- list.ggplotsbygroup$plots[[7]] + coord_flip()
TG7
```

# Chain Length Relation in TG's

_more background info in Literature Review_

Literature has stated that soil composition and lipid composition has been shown to adapt to environmental change including temperature. One major strategy by which *plants* adapt to temperature change is to decrease the degree of unsaturation of membrane lipids (aka increase saturation) under high temperature and increase [unsaturation] under low temperature. 

```{r TG dataframe for chain length and degree of saturation}
TGchain <- summary_class[c(23:31), c(1,4)] %>%
  separate(row.identity, c("length", "dbond_number"), ":") %>%
  mutate(length = str_replace(length, "TG\\(", "")) %>%
  mutate(dbond_number = str_replace(dbond_number, "\\)", "")) %>%
  mutate(dbond_number = str_replace(dbond_number, "_B", ""))
```
```{r TG plot of chain length and degree of saturation with standardized count}
TGplot <- ggplot(TGchain, aes(x=length, y=Standardized_Difference_Control_minus_Heated, size=dbond_number, color=dbond_number)) + geom_point() + ggtitle("Triacylglyerols") + ylab("Difference [C - H]")
TGplot
```

TG is the largest group (out of 7), and there was no distinguishable trend between saturation or chain length with abundance. If anything, the most saturated (48:1) has a high Heated Plot count over Control  while the least saturated (58:9) has a high Control Plot count over Heated. This is the opposite to findings in plants.  

</center>
