---
title: "Supplemental Lipid Figures"
author: "Hanna Choi"
date: "11/17/2020"
output:
  html_document:
    code_folding: hide
editor_options: 
  chunk_output_type: console
---
<center>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
library(readr)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(DT)
library(ggpubr)
```
# Making usable data for ggplot

```{r data with counts}
POS <- read.csv("Blanchard_Lipid_POS.csv", header=TRUE)
NEG <- read.csv("Blanchard_Lipid_NEG.csv", header=TRUE)
names(POS) <- gsub(x = names(POS), pattern = "_POS", replacement = "")  
names(NEG) <- gsub(x=names(POS), pattern="_NEG", replacement="")
POS$charge <- "POS"
NEG$charge <- "NEG"
data <- rbind(POS, NEG) # lipid data with counts

summary_class <- read_csv("significantlipidswithformula.csv")[-1] # made in comparealigned .Rmd
summary_class <- summary_class %>%
  rename(Main_Class = `Main Class`) #getting rid of the space in column
summary_class <- summary_class[order(summary_class$Main_Class, summary_class$Standardized_Difference_Control_minus_Heated),] #ordering the metadata

df <- merge(summary_class, data, by="row.identity", all=FALSE)
# For the dflong below:
# columns 17-43 is counts 
# Blanch_Nat_Lip is redundant
# separate by treatment Control vs Heated
dflong <- df[-c(11:16)] %>% 
  gather(plot, count,Blanch_Nat_Lip_C_12_AB_M_17:Blanch_Nat_Lip_H_4_AB_M_05) %>% 
  mutate(plot = gsub("Blanch_Nat_Lip_", "", plot)) %>% 
  separate(plot, c("treatment","plot"), "_", extra = "merge") %>%
  unite("row.id", c(row.identity, soil_type), remove=FALSE) %>%
  mutate(row.id = gsub("_Mineral", "", row.id)) %>%
  mutate(row.id = gsub("_Organic", "", row.id)) %>%
  arrange(soil_type)
dflongMineral <- dflong[which(dflong$soil_type == "Mineral"),]
dflongOrganic <- dflong[which(dflong$soil_type == "Organic"),]
dflong$CoverH <- dflong$Control_Average/dflong$Heated_Average

data$category <- gsub("\\(.*", "", data$row.identity)
dfdf <- data %>% gather(plot, count, Blanch_Nat_Lip_C_12_AB_M_17:Blanch_Nat_Lip_H_4_AB_M_05) %>%
  mutate(plot=gsub("Blanch_Nat_Lip_", "", plot)) %>%
  mutate(plot=gsub("[0-9]|_|A|B", "", plot)) %>%
  mutate(plot=factor(plot), category=factor(category)) %>%
  mutate(treatment=gsub("O|M", "",plot)) %>%
  mutate(plot=gsub("H|C", "", plot)) %>%
  unite(lipid, c(treatment, category), remove=FALSE) %>%
  mutate(lipid=factor(lipid))
#t <- dfdf %>%
#  group_by(category, lipid) %>%
#  arrange(category, lipid, desc(count)) %>%
#  mutate(mx = max(count))

d <- dfdf %>% group_by(row.identity,plot,treatment) %>%
  summarise(average =mean(count))
#write.csv(d, "Lipid-averages.csv")
testcon <- dfdf %>% group_by(row.identity,plot,treatment) %>%
  summarise(average =mean(count)) %>% filter(plot=="M", treatment=="C") %>% arrange(desc(average))
testcon$row.identity <- factor(testcon$row.identity)
levelorder <- testcon$row.identity
testconOrganic <- dfdf %>% group_by(row.identity,plot,treatment) %>%
  summarise(average =mean(count)) %>% filter(plot=="O", treatment=="C") %>% arrange(desc(average))
testconOrganic$row.identity <- factor(testconOrganic$row.identity)
levelorderOrganic <- testconOrganic$row.identity
#significantlipidswithformula <- read_csv("Lipidomics/significantlipidswithformula.csv")
#write.csv(levelorder, "significanceorderorganic.csv")
``` 
```{r}
significantlipidswithformula$P <- "P"
significantlipidswithformula$test <- significantlipidswithformula$row.identity
d.with.signif <- left_join(d,significantlipidswithformula[,c(2,16,17)])
d.with.signif[is.na(d.with.signif)] <- "N"
d.with.signif$test <- gsub("N", "", d.with.signif$test)
d.with.signif$test <- gsub("N", "", d.with.signif$test)
```
https://stackoverflow.com/questions/63082048/how-can-i-add-asterisks-for-the-significant-p-values-in-ggplot
```{r}
#write.csv(d.with.signif, "MineralHorizon.Lipidfigure.data.csv")
#Actually both Horizons :
MineralHorizon_Lipidfigure_dataeditedp <- read_csv("MineralHorizon.Lipidfigure.data.csv")
```
```{r}
Min <- MineralHorizon_Lipidfigure_dataeditedp[which(MineralHorizon_Lipidfigure_dataeditedp$plot=="M"),]
Min.Figure <- ggplot(MineralHorizon_Lipidfigure_dataeditedp[which(MineralHorizon_Lipidfigure_dataeditedp$plot=="M"),], aes(x=factor(row.identity, level=levelorder), y=log(average), color=treatment)) + 
  geom_point() + 
  geom_text(data=Min, aes(label = test), hjust = -1.5, vjust=-3, size=2) +
  coord_flip() +
  theme(axis.text=element_text(size=5), axis.title.y=element_blank(), axis.title.x=element_blank(), legend.position="none") + ggtitle("A. Mineral Horizon")
ggsave("Min.Figure.png")
```

```{r}
Org <- MineralHorizon_Lipidfigure_dataeditedp[which(MineralHorizon_Lipidfigure_dataeditedp$plot=="O"),]

Org.Figure <- ggplot(Org, aes(x=factor(row.identity, level=levelorderOrganic), y=log(average), color=treatment)) + 
  geom_point() + 
  geom_text(data=Org, aes(label = test), hjust = -1.5, vjust=-3, size=1) +
  coord_flip() +
  theme(axis.text=element_text(size=5), axis.title.y=element_blank(), axis.title.x=element_blank(), legend.position="none") + ggtitle("B. Organic Horizon")
ggsave("Org.Figure.png")
```

# Overarching Lipid Figure
```{r slightly messy but all significant lipids ordered by class, fig.cap = "Lipidomic data found 31 lipids that were altered in abundance significantly. 1 of the 31 lipid abundances were found to have increased due to an increase in temperature. From the figure, TG is Triacylglyerol"}
lipidfigureMineral <- ggplot(data=dflongMineral, aes(x=row.id, y=log(count), group_by(Main_Class, treatment), fill = treatment)) + geom_boxplot(show.legend = FALSE) + coord_flip() + scale_fill_manual(values = c( "#082BEA", "#EA0D08")) + scale_color_manual(values = c("#D4E126", "#12EF3A")) + ggtitle("Significant Lipids Heated v. Control Soil Mineral Plots") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
lipidfigureMineral
lipidfigureOrganic <- ggplot(data=dflongOrganic, aes(x=row.id, y=log(count), group_by(Main_Class, treatment), fill = treatment)) + geom_boxplot() + coord_flip() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + scale_fill_manual(values = c( "#082BEA", "#EA0D08")) + scale_color_manual(values = c("#D4E126", "#12EF3A")) + ggtitle("Significant Lipids Heated v. Control Soil Organic Plots")
lipidfigure <- ggarrange(lipidfigureMineral, lipidfigureOrganic, ncol=1)
lipidfigure
```
# Making plots by Class

```{r ggplots}
list.ggplotsbygroup <- dflong %>% group_by(Main_Class) %>% do(plots=ggplot(data=.) + aes(x=row.id, y=log(count)) + geom_boxplot(aes(fill=treatment)) + xlab("compound") + ylab("abundance (logged)") + ggtitle(unique(.$Main_Class)))
```
# Analysis by Group
```{r, fig.cap="DG; inconclusive due to number of compounds being low and no significant differences between chain lengths or degrees of saturation"}
DG1 <- list.ggplotsbygroup$plots[[1]] + coord_flip()
DG1
```

---

```{r, fig.cap = "PC; there's an interesting difference between PC_A and PC_B which only differ in stereochemistry"}
PC2 <- list.ggplotsbygroup$plots[[2]] + coord_flip()
PC2
```

---

```{r, fig.cap = "PE is found to be a plant growth regulator"}
PE3 <- list.ggplotsbygroup$plots[[3]] + coord_flip()
PE3
```

---

```{r, fig.cap = "DGDG; decreased abundance for all compunds"}
DGDG4 <- list.ggplotsbygroup$plots[[4]] + coord_flip()
DGDG4
```

---

```{r, fig.cap = "Only 1 hexosylceramides with two different stereochemical forms found to be significantly altered due to Temeprature. The abundance signficantly decreased in the Heated plots"}
HexCer5 <- list.ggplotsbygroup$plots[[5]] + coord_flip()
HexCer5
```

---

```{r, fig.cap = "Significant DGTSA's decreased in abundance as well"}
DGTSA6 <- list.ggplotsbygroup$plots[[6]] + coord_flip()
DGTSA6
```

---

```{r, fig.cap = "TG's are most common lipids in soil and TG(58:9) was found to be the only lipid to increase in abundance in the heated plots over the control plots"}
TG7 <- list.ggplotsbygroup$plots[[7]] + coord_flip()
TG7
```

# Chain Length Relation in TG's

_more background info in Literature Review_

Literature has stated that soil composition and lipid composition has been shown to adapt to environmental change including temperature. One major strategy by which *plants* adapt to temperature change is to decrease the degree of unsaturation of membrane lipids (aka increase saturation) under high temperature and increase [unsaturation] under low temperature. 

```{r TG dataframe for chain length and degree of saturation}
TGchain <- summary_class[c(23:31), c(1,4)] %>%
  separate(row.identity, c("length", "dbond_number"), ":") %>%
  mutate(length = str_replace(length, "TG\\(", "")) %>%
  mutate(dbond_number = str_replace(dbond_number, "\\)", "")) %>%
  mutate(dbond_number = str_replace(dbond_number, "_B", ""))
```
```{r TG plot of chain length and degree of saturation with standardized count}
TGplot <- ggplot(TGchain, aes(x=length, y=Standardized_Difference_Control_minus_Heated, size=dbond_number, color=dbond_number)) + geom_point() + ggtitle("Triacylglyerols") + ylab("Difference [C - H]")
TGplot
```

TG is the largest group (out of 7), and there was no distinguishable trend between saturation or chain length with abundance. If anything, the most saturated (48:1) has a high Heated Plot count over Control  while the least saturated (58:9) has a high Control Plot count over Heated. This is the opposite to findings in plants.  

</center>
