---
title: "omu using KEGGmatches"
author: "Hanna Choi"
output: html_document
---

## content

My count data is countdatacsv; df is the count data with categories. mymetadata is my metadata. I am trying to put the df and mymetedata into omu_summary, a function that gives t-test statistics and even folding analysis. I am following this vignette:
https://cran.r-project.org/web/packages/omu/vignettes/Omu_vignette.html


```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE)
library(omu)
library(readr)
library(tidyverse)
library(DT)
library(rstatix)
library(ggpubr)
countdata <- read_metabo("EMSL_PNNL_plusKEGG.csv")[-1]
# reads as dataframe not 'array'
#first column was just indices so they were deleted

countdatacsv <- read_csv("EMSL_PNNL_plusKEGG.csv")[-1]
# using read_csv because read_metabo doesn't work for mutate() 
countdatacsv <- countdatacsv %>% mutate_if(is.numeric, round, 0) %>%
  mutate_if(is.double,as.integer) %>%
  mutate_if(is.character,as.factor)

mymetadata <- read_csv("EMSL_PNNL_metadata.csv")[-1] #ditto
mymetadata <- mymetadata %>%
  mutate(Grouped = gsub("_", "", Grouped)) %>%
  mutate_if(is.character, as.factor)
head(mymetadata)
```

```{r their example data analysis}
# their example (below) works fine
# omu package included c57_nos2KO_mouse_metadata and c57_nos2KO_mouse_countDF
# assigning hierarchy characterizes the metabolites:
DF <- assign_hierarchy(count_data = c57_nos2KO_mouse_countDF, keep_unknowns = TRUE, identifier = "KEGG")
theirsummary <- omu_summary(count_data = DF, metadata = c57_nos2KO_mouse_metadata, numerator = "Strep", denominator = "Mock", response_variable = "Metabolite", Factor = "Treatment", log_transform = TRUE, p_adjust = "BH", test_type = "welch")
```

### Assigning omu hierarchical class data using the KEGG Compound Name

```{r}
df <- assign_hierarchy(count_data = countdata, keep_unknowns = TRUE, identifier = "KEGG")
```

### 'omu_summary' intro for summary statistical data using data and metadata; this doesn't work.

> I've tried changing the type of values in the df columns from character to factor and double to integer to mirror their example data

```{r eval = FALSE}
df_stats <- omu_summary(count_data = df, metadata = mymetadata, numerator = "H", denominator = "C", response_variable = "Metabolite", Factor = "Treatment", log_transform = FALSE, p_adjust = "BH", test_type = "welch")
# if csv version of data is read, Error in `.rowNamesDF<-`(x, value = value) : invalid 'row.names' length
```

### manually getting stats

```{r}
cleancountdata <- countdata %>% gather(key = "plot", value = "count", M_C_4_AB:O_H_32_AB) %>% mutate(type = gsub("_[0-9][0-9]*_AB","",plot)) %>% mutate(Min.Org = gsub("_[A-Z]","",type)) %>% unite(metabplot, c("Metabolite","Min.Org")) %>% mutate(metabplot = gsub("_2B", "", metabplot)) %>% mutate(type = gsub("_2B","", type))

cleancountdata[cleancountdata == 0] <- 1
cleancountdata$count <- log(cleancountdata$count)
metab.stat <- cleancountdata %>%
  group_by(metabplot) %>%
  t_test(count ~ type) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance()
# summary; mean and st.dev:
MCmean <- cleancountdata %>% group_by(metabplot, type) %>% filter(type == "M_C") %>%
  summarise(meanC = mean(count), st.devC = sd(count)) %>% select(-type)
MHmean <- cleancountdata %>% group_by(metabplot, type) %>% filter(type == "M_H") %>%
  summarise(meanH = mean(count), st.devH = sd(count)) %>% select(-type)
OCmean <- cleancountdata %>% group_by(metabplot, type) %>% filter(type == "O_C") %>%
  summarise(meanC = mean(count), st.devC = sd(count)) %>% select(-type)
OHmean <- cleancountdata %>% group_by(metabplot, type) %>% filter(type == "O_H") %>%
  summarise(meanH = mean(count), st.devH = sd(count)) %>% select(-type)
means <- rbind(merge(MCmean, MHmean), merge(OCmean,OHmean)) 
rm(MCmean, MHmean, OCmean,OHmean)
# combine summary and stats:
metab.stat.sum <- merge(metab.stat, means, by = "metabplot", all=TRUE)
metab.stat.sum$foldchange <- metab.stat.sum$meanC/metab.stat.sum$meanH
metab.stat.sum$log2foldchange <- log2(metab.stat.sum$foldchange)
metab.stat.sum$Metabolite <- gsub("_[A-Z]","",metab.stat.sum$metabplot)
# KEGG and heirarchy matching:
KEGG_ID <- read_csv("KEGG.ID.csv")
df <- merge(metab.stat.sum, KEGG_ID, by = "Metabolite", all = TRUE)
df <- as.data.frame(assign_hierarchy(df, keep_unknowns = TRUE, identifier = "KEGG"))
sig <- df[which(df$p <= 0.05),]
sig[1]
df(which(df$Metabolite == "1,3-dihydroxyacetone"))
# just significant metabolites
# write.csv(df, "Metabolites.withstat.KEGG.hierarchy.foldchange.csv")
```

I found `r nrow(sig)` deferentially significant metabolites in both layers with `r sum(sig$group1 == "O_C")` being in the Organic layer while `r sum(sig$group1 == "M_C")` being in the Mineral layer.

### fold change

```{r}
ggplot(metab.stat.sum) + geom_point(aes(x=log2foldchange, y=-log(p.adj), color = p.adj.signif)) + theme(legend.position = "none",
              plot.title = element_text(size = rel(1.5), hjust = 0.5),
              axis.title = element_text(size = rel(1.25)))
```

### To Do: look at slide 20 from AGU presentation and make similar image for metabolites

```{r}
df$meancountdifference <- df$meanH - df$meanC
makeStars <- function(x){
  stars <- c("*", "ns")
  vec <- c(0, 0.05 )
  i <- findInterval(x, vec)
  stars[i]
}
df$p.category <- makeStars(df$p)
```

```{r}
x <- df[which(df$Metabolite %in% sig$Metabolite),]
# df %>% mutate(group1 = replace(group1, val = "M_C", "Mineral")) %>% mutate(group1 = replace(group1, val = "O_C", "Organic"))
significantmetabolites <- ggplot(df[which(df$Metabolite %in% sig$Metabolite),], aes(x=Metabolite, y=meancountdifference, color = p.category)) + geom_point() + theme(axis.text.y = element_text(hjust = 0.85)) + ylab("average count difference (H - C)") + xlab("Metabolite") + coord_flip() + ggtitle("Significant Metabolites") + facet_grid(~group1)
mutate(df[,4], group1 = ifelse(group1 == "M_C", "Mineral", group1))
significantmetabolites
```


